---
title: "Week 3 - tmap"
format: 
  revealjs:
    self-contained: true
execute: 
  echo: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(sf)
library(stars)

options(width=70)
```

1 Rapa Nui / Isla de Pascua ((668715.4 7002628, 668776.6 7002640, 668796

## Data

The example data is taken from the [Elegant and informative maps with tmap
](https://r-tmap.github.io/tmap-book/) book - see the data folder in the github repository.

Filetypes:

* `tif` files are geospatial raster data (e.g. elevation maps)

* `gpkg` are geopackage files (modern version of shapefiles)


## Easter Island data - elevation

```{r}
(elev = read_stars("data/easter_island/ei_elev.tif"))
```

##

```{r}
plot(elev)
```

## Easter Island data - border

```{r}
(border = read_sf("data/easter_island/ei_border.gpkg"))
```

##

```{r}
plot(border)
```


## Easter Island data - roads

```{r}
(roads = read_sf("data/easter_island/ei_roads.gpkg"))
```

##

```{r}
plot(roads %>% select(type))
```


## Easter Island data - points of interest

```{r}
(points = read_sf("data/easter_island/ei_points.gpkg"))
```

##

```{r}
plot(points %>% select(type), pch=16)
```


## Layers with ggplot

```{r ggmap, eval=FALSE}
ggplot() +
  geom_sf(data=border) +
  scale_fill_distiller(palette="RdYlGn", na.value = "transparent") +
  geom_stars(data=elev) +
  geom_sf(data=roads, aes(linewidth=strokelwd)) +
  scale_linewidth_continuous(range = c(0.1,1)) +
  geom_sf(data=points, aes(color=type, size=elevation), shape=17) +
  scale_size_continuous(range=c(1,4)) +
  theme_minimal() +
  guides(linewidth="none")
```

##

```{r ref.label="ggmap", echo=FALSE}
```


# tmap

## tmap

> tmap is an actively maintained open-source R-library for drawing thematic maps. The API is based on A Layered Grammar of Graphics and resembles the syntax of ggplot2, a popular R-library for drawing charts.

```{r}
library(tmap)
```

## Easter Island - Border

```{r}
#| output-location: slide
tm_shape(border) +
  tm_borders()
```

## 

```{r}
#| output-location: slide
tm_shape(border) +
  tm_borders() +
  tm_layout(
    main.title = "Easter Island",
    bg.color="lightblue"
  )
```

## Roads

```{r}
#| output-location: slide
tm_shape(border) +
  tm_borders() +
tm_shape(roads) +
  tm_lines(lwd="strokelwd", legend.lwd.show = FALSE, ) +
tm_layout(
  main.title = "Easter Island",
  bg.color="lightblue"
)
```

##

```{r}
#| output-location: slide
tm_shape(border) +
  tm_borders() +
tm_shape(roads) +
  tm_lines(
    lwd="strokelwd", 
    legend.lwd.show = FALSE, 
    col = "type"
  ) +
tm_layout(
  main.title = "Easter Island",
  bg.color="lightblue"
)
```

## Elevation

```{r}
#| output-location: slide
tm_shape(elev) +
  tm_raster() +
tm_shape(border) +
  tm_borders() +
tm_shape(roads) +
  tm_lines(
    lwd="strokelwd", 
    legend.lwd.show = FALSE
  ) +
tm_layout(
  main.title = "Easter Island",
  bg.color="lightblue"
)
```


## 

```{r}
#| output-location: slide
tm_shape(elev) +
  tm_raster(
    style = "cont",
    title = "Elevation (m)"
  ) +
tm_shape(border) +
  tm_borders() +
tm_shape(roads) +
  tm_lines(
    lwd="strokelwd", 
    legend.lwd.show = FALSE
  ) +
tm_layout(
  main.title = "Easter Island",
  bg.color="lightblue"
)
```

## 

```{r}
#| output-location: slide
tm_shape(elev) +
  tm_raster(
    style = "cont",
    title = "Elevation (m)",
    palette = "-RdYlGn"
  ) +
tm_shape(border) +
  tm_borders() +
tm_shape(roads) +
  tm_lines(
    lwd="strokelwd", 
    legend.lwd.show = FALSE
  ) +
tm_layout(
  main.title = "Easter Island",
  bg.color="lightblue"
)
```

## Volcanoes {.smaller}

```{r}
#| output-location: slide
tm_shape(elev) +
  tm_raster(
    style = "cont",
    title = "Elevation (m)",
    palette = "-RdYlGn"
  ) +
tm_shape(border) +
  tm_borders() +
tm_shape(roads) +
  tm_lines(
    lwd="strokelwd", 
    legend.lwd.show = FALSE
  ) +
tm_shape(
  points %>% filter(type == "volcano")
) + 
  tm_symbols(
    shape = 24,
    size = "elevation",
    title.size = "Volcanoes (m)"
  ) +
tm_layout(
  main.title = "Easter Island",
  bg.color="lightblue"
)
```


## Other stuff {.smaller}

```{r}
#| output-location: slide
tm_graticules(alpha=0.25) +
tm_shape(elev) +
  tm_raster(
    style = "cont",
    title = "Elevation (m)",
    palette = "-RdYlGn"
  ) +
tm_shape(border) +
  tm_borders() +
tm_shape(roads) +
  tm_lines(
    lwd="strokelwd", 
    legend.lwd.show = FALSE
  ) +
tm_shape(
  points %>% filter(type == "volcano")
) + 
  tm_symbols(
    shape = 24,
    size = "elevation",
    title.size = "Volcanoes (m)"
  ) +
tm_layout(
  main.title = "Easter Island",
  bg.color="lightblue"
) +
tm_compass(position = c("right", "top")) +
tm_scale_bar() +
tm_credits("Author, 2021") +
tm_add_legend(type = "line", col = "black", title = "Roads")
```



## tmap modes

```{r}
map = tm_graticules(alpha=0.25) +
tm_shape(elev) +
  tm_raster(
    style = "cont",
    title = "Elevation (m)",
    palette = "-RdYlGn"
  ) +
tm_shape(border) +
  tm_borders() +
tm_shape(roads) +
  tm_lines(
    lwd="strokelwd", 
    legend.lwd.show = FALSE
  ) +
tm_shape(
  points %>% filter(type == "volcano")
) + 
  tm_symbols(
    shape = 24,
    size = "elevation",
    title.size = "Volcanoes (m)"
  ) +
tm_layout(
  main.title = "Easter Island",
  bg.color="lightblue"
) +
tm_compass(position = c("right", "top")) +
tm_scale_bar() +
tm_credits("Author, 2021") +
tm_add_legend(type = "line", col = "black", title = "Roads")
```

## plot

```{r}
tmap_mode("plot")
map
```

## view

```{r}
tmap_mode("view")
map
```

## Tiles

```{r}
#| output-location: slide
tm_basemap(
  c( osm  = "OpenStreetMap",
     esri = "Esri.WorldStreetMap")
) +
tm_shape(border, is.master = TRUE) +
  tm_borders() +
tm_shape(roads) +
  tm_lines(
    lwd="strokelwd", 
    legend.lwd.show = FALSE
  ) +
tm_shape(
  points %>% filter(type == "volcano")
) + 
  tm_symbols(
    size = "elevation",
    legend.size.show = FALSE
  )
```

## Tiles w/ plot mode

```{r}
#| output-location: slide
tmap_mode("plot")
tm_basemap(
  c( osm  = "OpenStreetMap",
     esri = "Esri.WorldStreetMap")
) +
tm_shape(border, is.master = TRUE) +
  tm_borders() +
tm_shape(roads) +
  tm_lines(
    lwd="strokelwd", 
    legend.lwd.show = FALSE
  ) +
tm_shape(
  points %>% filter(type == "volcano")
) + 
  tm_symbols(
    size = "elevation",
    legend.size.show = FALSE
  )
```

##

```{r}
#| output-location: slide
tmap_mode("plot")
tm_shape(
  tmaptools::read_osm(border)
) +
  tm_rgb() +
tm_shape(border, is.master = TRUE) +
  tm_borders() +
tm_shape(roads) +
  tm_lines(
    lwd="strokelwd", 
    legend.lwd.show = FALSE
  ) +
tm_shape(
  points %>% filter(type == "volcano")
) + 
  tm_symbols(
    size = "elevation",
    legend.size.show = FALSE
  )
```